### 2015-03-19 swmpr_manu.tex
started buffing up text

The next example illustrates how to handle missing values using the \texttt{decomp} function. The \texttt{decompose} function used internally within \texttt{decomp} currently cannot process time series with missing values.  A recommended approach is to use \texttt{na.approx} to interpolate the missing values prior to \texttt{decompose}.

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{kframe}
\begin{alltt}
\hlcom{# get data}
\hlstd{dat} \hlkwb{<-} \hlkwd{subset}\hlstd{(swmp1,} \hlkwc{subset} \hlstd{=} \hlkwd{c}\hlstd{(}\hlstr{'2013-06-01 00:00'}\hlstd{,} \hlstr{'2013-07-31 00:00'}\hlstd{))}

\hlcom{# this returns an error}
\hlcom{# test <- decomp(dat, param = 'do_mgl', frequency = 'daily')}

\hlcom{# how many missing values?}
\hlkwd{sum}\hlstd{(}\hlkwd{is.na}\hlstd{(dat}\hlopt{$}\hlstd{do_mgl))}
\end{alltt}
\begin{verbatim}
## [1] 3
\end{verbatim}
\begin{alltt}
\hlcom{# use na.approx to interpolate missing data}
\hlstd{dat} \hlkwb{<-} \hlkwd{na.approx}\hlstd{(dat,} \hlkwc{params} \hlstd{=} \hlstr{'do_mgl'}\hlstd{,} \hlkwc{maxgap} \hlstd{=} \hlnum{10}\hlstd{)}

@@

An alternative approach to time series decomposition is provided by the \texttt{decomp\_cj} function, which is a simple wrapper to the \texttt{decompTs} function in the wq package.  Theory describing this method is described in Cloern and Jassby (2010).  The function is similar to \texttt{decomp.swmpr} with a few key differences.  The \texttt{decomp.swmpr} function decomposes the time series into a trend, seasonal, and random component, whereas the current function decomposes into the grandmean, annual, seasonal, and events components.  For both functions, the random or events components, respectively, can be considered anomalies that don't follow the trends in the remaining categories.  The \texttt{decomp\_cj} function provides only a monthly decomposition, which is appropriate for characterizing relatively long-term trends.  This approach is meant for nutrient data that are obtained on a monthly cycle.  The function will also work with continuous water quality or weather data but note that the data are first aggregated on the monthly scale before decomposition.  Accordingly, short-term variation less than one-month will be removed. Additional arguments passed to \texttt{decompTs} can be used with \texttt{decomp\_cj}, such as \texttt{startyr}, \texttt{endyr}, and \texttt{type}.  Values passed to \texttt{type} are \texttt{mult} (default) or \texttt{add}, referring to multiplicative or additive decomposition.  See the documentation for \texttt{decompTs} for additional explanation and examples.

@@

A reserve map with all stations can be obtained using the \texttt{map\_reserve} function.  This function is a simple wrapper to functions in the ggmap package. The current function is limited to Google maps, which allows four map types that can be set with the \texttt{map\_type} argument: terrain (default), satellite, roadmap, or hybrid.  The \texttt{zoom} argument may have to be chosen through trial and error depending on the spatial extent of the reserve.  See the help documentation for ggmap for more info on zoom.  Additionally, station locations are returned using the \texttt{site\_codes\_ind} function if the computer making the request has the IP address registered with CDMO. Otherwise, a local and possibly outdated file is used.  Use the contact at the CDMO \href{http://cdmo.baruch.sc.edu/webservices.cfm}{web services} to register your IP.

{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-201}

\begin{kframe}\begin{alltt}
\hlcom{# plot the stations at Padilla Bay reserve, satellite}
\hlkwd{map_reserve}\hlstd{(}\hlstr{'pdb'}\hlstd{,} \hlkwc{map_type} \hlstd{=} \hlstr{'satellite'}\hlstd{,} \hlkwc{zoom} \hlstd{=} \hlnum{12}\hlstd{)}
\end{alltt}
\end{kframe}

{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-202}

@@

Several graphics showing seasonal and annual trends for a given SWMP parameter can be obtained using the \texttt{plot\_summary} function.  The plots include monthly distributions, monthly anomalies, and annual anomalies in multiple formats.  Anomalies are defined as the difference between the monthly or annual average from the grand mean for the parameter.  Monthly anomalies are in relation to the grand mean for the same month across all years.  All data are aggregated for quicker plotting.  Nutrient data are based on monthly averages, whereas weather and water quality data are based on daily averages.  Cumulative precipitation data are based on the daily maximum. The function returns a graphics object (Grob) of multiple ggplot objects.  An interactive Shiny widget that uses this function is available: \href{https://beckmw.shinyapps.io/swmp_summary/}{https://beckmw.shinyapps.io/swmp\_summary/}.

Estimates of ecosystem metabolism provide a useful measure of overall system productivity.  These estimates are commonly used to evaluate whether an ecosystem is a net source or sink of organic material.  The open-water method is a common approach to quantify net ecosystem metabolism using a mass balance equation that describes the change in dissolved oxygen over time from the balance between photosynthetic and respiration rates, corrected for air-sea gas diffusion at the surface.  The diffusion-corrected DO flux estimates are averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of DO change. DO flux is averaged during night hours for respiration and averaged during day hours for net production. Respiration rates are assumed constant during day and night such that total daily rates are calculated as hourly respiration multiplied by 24. The metabolic day is considered the 24 hour period between sunsets on two adjacent calendar days.  Respiration is subtracted from daily net production estimates to yield gross production.

The \texttt{ecometab} function is used to implement an adaptation of the open-water method.  Several assumptions must be met for a valid interpretation of the results.  In general, the dissolved oxygen time series is assumed to represent the same water mass over time.  Tidal advection may have a significant influence on the time series, which can contribute to a significant amount of noise in metabolic estimates.  The extent to which tidal advection influences the dissolved oxygen signal depends on various site-level characteristics and an intimate knowledge of the site may be required.  Volumetric rates for gross production and total respiration are also based on total depth of the water column, which is assumed to be mixed.  Water column depth is based on mean value for the depth variable across the time series and is floored at 1 meter for very shallow stations.  Additionally, the volumetric reaeration coefficient requires an estimate of the anemometer height of the weather station, which is set as 10 meters by default.  The metadata should be consulted for exact height. Other assumptions may apply and the user should consult the relevant literature (see the references in the help file).  All estimates are in mmol of oxygen but can be converted  to grams by changing the default arguments (i.e., 1mmol O2 = 32 mg O2, 1000 mg = 1g, multiply all estimates by 32/1000).

@@

{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-22}

>>>>

An alternative approach to time series decomposition is provided by the \texttt{decomp\_cj} function, which is a simple wrapper to the \texttt{decompTs} function in the wq package.  Theory describing this method is described in \cite{Cloern10}.  The function is similar to \texttt{decomp.swmpr} with a few key differences.  The \texttt{decomp.swmpr} function decomposes the time series into a trend, seasonal, and random component, whereas the current function decomposes into the grandmean, annual, seasonal, and events components.  For both functions, the random or events components, respectively, can be considered anomalies that don't follow the trends in the remaining categories.  The \texttt{decomp\_cj} function provides only a monthly decomposition, which is appropriate for characterizing relatively long-term trends.  This approach is meant for nutrient data that are obtained on a monthly cycle.  The function will also work with continuous water quality or weather data but note that the data are first aggregated on the monthly scale before decomposition.  Accordingly, short-term variation less than one-month will be removed. Additional arguments passed to \texttt{decompTs} can be used with \texttt{decomp\_cj}, such as \texttt{startyr}, \texttt{endyr}, and \texttt{type}.  Values passed to \texttt{type} are \texttt{mult} (default) or \texttt{add}, referring to multiplicative or additive decomposition.  See the documentation for \texttt{decompTs} for additional explanation and examples.

@@

A reserve map with all stations can be obtained using the \texttt{map\_reserve} function.  This function is a simple wrapper to functions in the ggmap package. The current function is limited to Google maps, which allows four map types that can be set with the \texttt{map\_type} argument: terrain (default), satellite, roadmap, or hybrid.  The \texttt{zoom} argument may have to be chosen through trial and error depending on the spatial extent of the reserve.  See the help documentation for ggmap for more info on zoom.  Additionally, station locations are returned using the \texttt{site\_codes\_ind} function if the computer making the request has the IP address registered with \ac{CDMO}. Otherwise, a local and possibly outdated file is used.  Use the contact at the \ac{CDMO} \href{http://cdmo.baruch.sc.edu/webservices.cfm}{web services} to register your IP.

@@

Several graphics showing seasonal and annual trends for a given SWMP parameter can be obtained using the \texttt{plot\_summary} function.  The plots include monthly distributions, monthly anomalies, and annual anomalies in multiple formats.  Anomalies are defined as the difference between the monthly or annual average from the grand mean for the parameter.  Monthly anomalies are in relation to the grand mean for the same month across all years.  All data are aggregated for quicker plotting.  Nutrient data are based on monthly averages, whereas weather and water quality data are based on daily averages.  Cumulative precipitation data are based on the daily maximum. The function returns a graphics object (Grob) of multiple ggplot objects.  An interactive Shiny widget that uses this function is available (see the \nameref{supp_info}).

{\centering \includegraphics[width=\maxwidth]{figure/unnamed-chunk-20}

@@

Estimates of ecosystem metabolism provide a useful measure of overall system productivity.  These estimates are commonly used to evaluate whether an ecosystem is a net source or sink of organic material.  The open-water method is a common approach to quantify net ecosystem metabolism using a mass balance equation that describes the change in dissolved oxygen over time from the balance between photosynthetic and respiration rates, corrected for air-sea gas diffusion at the surface \cite{Odum56}.  The diffusion-corrected DO flux estimates are averaged during day and night for each 24 hour period in the time series, where flux is an hourly rate of DO change. DO flux is averaged during night hours for respiration and averaged during day hours for net production. Respiration rates are assumed constant during day and night such that total daily rates are calculated as hourly respiration multiplied by 24. The metabolic day is considered the 24 hour period between sunsets on two adjacent calendar days.  Respiration is subtracted from daily net production estimates to yield gross production.

The \texttt{ecometab} function is used to implement an adaptation of the open-water method \cite{Odum56,Caffrey14}.  Several assumptions must be met for a valid interpretation of the results.  In general, the dissolved oxygen time series is assumed to represent the same water mass over time.  Tidal advection may have a significant influence on the time series, which can contribute to a significant amount of noise in metabolic estimates.  The extent to which tidal advection influences the dissolved oxygen signal depends on various site-level characteristics and an intimate knowledge of the site may be required.  Volumetric rates for gross production and total respiration are also based on total depth of the water column, which is assumed to be mixed.  Water column depth is based on mean value for the depth variable across the time series and is floored at 1 meter for very shallow stations.  Additionally, the volumetric reaeration coefficient requires an estimate of the anemometer height of the weather station, which is set as 10 meters by default.  The metadata should be consulted for exact height. Other assumptions may apply and the user should consult the relevant literature (see the references in the help file).  All estimates are in mmol of oxygen but can be converted  to grams by changing the default arguments (i.e., 1mmol O2 = 32 mg O2, 1000 mg = 1g, multiply all estimates by 32/1000).

@@
